---
title: "Multinomial Logistic Regression Tutorial"
author: "Barbara Flores"
format:
  pdf:
    toc: true
    number-sections: true
    number-depth: 3
output: 
  pdf_document: 
    latex_engine: xelatex
    keep_tex: true
    latex_args: ["-shell-escape"]
fontsize: 11pt
header-includes:
  - "\\usepackage[left=0.75in,right=0.75in,top=0.75in,bottom=1.25in]{geometry}"
  - "\\usepackage{enumitem}"
  - "\\setlist{noitemsep, topsep=0pt, parsep=2pt, partopsep=0pt}"
  - "\\setlength{\\parskip}{0pt}"
  - "\\RedeclareSectionCommand[beforeskip=-1sp,afterskip=0.5\\baselineskip]{section}"
  - "\\RedeclareSectionCommand[beforeskip=-1sp,afterskip=0.25\\baselineskip]{subsection}"

editor: 
  markdown: 
    wrap: 72
---

\setcounter{page}{0}

\thispagestyle{empty}

{{< pagebreak >}}

```{r, include = FALSE}
# Global settings to hide code and results in the final document
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results= 'hide')
```

```{r}
library(pacman)
p_load(openintro, tidyverse,dplyr,tidyr, caret, pROC, e1071, sjPlot)
```

```{r}
path = "https://raw.githubusercontent.com/BarbaraPFloresRios/IDS702_ModelingAndRepresentationOfData/main/20231117_MultinomialLogisticRegression/MultinomialData.csv"

MultinomialData <-  read_csv(path)
glimpse(MultinomialData)

```

# Overview

## Generalized linear model

\vspace{0.3cm}

Generalized Linear Models (GLMs) are a class of statistical models in
which the relationship between a dependent variable and one or more
independent variables follows a specific probability distribution. The
general formula of a Generalized Linear Model (GLM) is expressed as
follows:

$$
g(μ) = Xβ
$$

where

\- $g(μ)$: The link function connecting the mean $μ$ of the response
variable distribution with the linear combination of predictor variables
$Xβ$

\- $X$:The design matrix containing the values of the predictor
variables.

\- $β$: The vector of coefficients associated with each predictor
variable.

The choice of probability distribution and the link function allows the
adaptation of the model to different types of data and relationships
between variables."

### Link function

-   The link function connects the distribution of the outcome to Y Xβ
    and indicates how the expected value of the response relates to the
    linear combination of explanatory variables.

-   GLM assumptions include independent distribution of data, a
    specified distribution for the dependent variable (e.g., Bernoulli
    for logistic regression), and a linear relationship between the
    transformed expected response and the explanatory variables.

```{=html}
<!-- -->
```
-   Logistic regression and link functions

-   Motivation for using logistic regression

-   Outcome variable type

-   Model example data on pumpkin seeds

-   GLM terms: generalized linear models and link functions

-   GLM assumptions: independent data, specified distribution for
    dependent variable, linear relationship between transformed expected
    response and explanatory variables

-   Explanation of why there is no error term

```{=html}
<!-- -->
```
-   Motivation for logistic regression and link functions

-   Logistic regression is used for binary outcome variables

-   Generalized linear models (GLMs) assume the response variable
    follows a distribution

-   Link function connects the distribution of the outcome to the linear
    combination of explanatory variables

-   GLM assumptions include independent distribution of data, specified
    distribution for the dependent variable, and linear relationship
    between the transformed expected response and explanatory variables

### Link function

### GLM Assumptions

## Multinomial Logistic Regression

$Pr(\theta | y) \propto Pr(y | \theta) Pr(\theta)$

$x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$

# Probability Distribution

# Data Example:

# Probability Distribution
