---
title: "Financial Independence Survey - Logistic Regression Model"
author: "Barbara Flores"
format:
  pdf:
    toc: true
    number-sections: true
    number-depth: 3
output: 
  pdf_document: 
    latex_engine: xelatex
    keep_tex: true
    latex_args: ["-shell-escape"]
fontsize: 11pt
header-includes:
  - "\\usepackage[left=0.75in,right=0.75in,top=0.75in,bottom=1.25in]{geometry}"
  - "\\usepackage{enumitem}"
  - "\\setlist{noitemsep, topsep=0pt, parsep=2pt, partopsep=0pt}"
  - "\\setlength{\\parskip}{0pt}"
  - "\\RedeclareSectionCommand[beforeskip=-1sp,afterskip=0.5\\baselineskip]{section}"
  - "\\RedeclareSectionCommand[beforeskip=-1sp,afterskip=0.25\\baselineskip]{subsection}"

editor: 
  markdown: 
    wrap: 72
---

\setcounter{page}{0}

\thispagestyle{empty}

{{< pagebreak >}}

```{r, include = FALSE}
# Global settings to hide code and results in the final document
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results= 'hide')
```

```{r}
library(pacman)
p_load(openintro, tidyverse,dplyr,tidyr, caret, pROC, e1071, sjPlot)
```

```{r}
 data("reddit_finance")
```

```{r}
# Cleaning data and selecting relevant columns for predictive modeling
# First, We created the reddit_finance dataset for easier work. 
reddit_finance_sub <- reddit_finance 

# Then, we create new variables that allow us to use the current variables in our model.

# We will make some assumptions and decision rules based on criteria to clean our dataset, which we will detail in the following code:

# 1. num_incomes: 
# How many individuals contribute to your household income?
# We have 1 record with NA, which we will replace with 0, and 2 records with n > 3, which we will replace with 4.


reddit_finance_sub$num_incomes <- ifelse(is.na(reddit_finance_sub$num_incomes), 0, reddit_finance_sub$num_incomes)

reddit_finance_sub$num_incomes <- ifelse(reddit_finance_sub$num_incomes == ">3", 4, reddit_finance_sub$num_incomes)


reddit_finance_sub$num_incomes <- as.numeric(reddit_finance_sub$num_incomes)

# 2. pan_inc_chg
# As a result of the pandemic, did your earned income increase, decrease, or remain the same?

reddit_finance_sub$pan_inc_chg <- factor(reddit_finance_sub$pan_inc_chg, levels = c( "Decreased","Stayed the same","Increased"))

# 3. pan_inc_chg_pct
# By how much did your earned income change?
# For this variable, we apply the decision rule of replacing the intervals with their order number. For example, the interval 1-10% will be replaced by 1, 11-20% by 2, and so on.

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(pan_inc_chg_pct = case_when(
    pan_inc_chg_pct == "1-10%"  ~ 1,
    pan_inc_chg_pct == "11-20%" ~ 2,
    pan_inc_chg_pct == "21-30%" ~ 3,
    pan_inc_chg_pct == "31-40%" ~ 4,
    pan_inc_chg_pct == "41-50%" ~ 5,
    pan_inc_chg_pct == "51-60%" ~ 6,
    pan_inc_chg_pct == "61-70%" ~ 7,
    pan_inc_chg_pct == "71-80%" ~ 8,
    pan_inc_chg_pct == "81-90%" ~ 9,
    pan_inc_chg_pct == "91-100%" ~ 10,
    pan_inc_chg_pct == "No change" ~ 0,
    TRUE ~ as.numeric(pan_inc_chg_pct)  
  ))


reddit_finance_sub$pan_inc_chg_pct <- as.numeric(reddit_finance_sub$pan_inc_chg_pct)

# 4.pan_exp_chg
# As a result of the pandemic, did your expenses increase, decrease, or remain the same?

reddit_finance_sub$pan_exp_chg <- factor(reddit_finance_sub$pan_exp_chg, levels = c( "Decreased","Stayed the same", "Increased"))

# 5.pan_exp_chg_pct
# By how much did your expenses change?
# For this variable, we apply the decision rule of replacing the intervals with their order number. For example, the interval 1-10% will be replaced by 1, 11-20% by 2, and so on.

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(pan_exp_chg_pct = case_when(
    pan_exp_chg_pct == "1-10%"  ~ 1,
    pan_exp_chg_pct == "11-20%" ~ 2,
    pan_exp_chg_pct == "21-30%" ~ 3,
    pan_exp_chg_pct == "31-40%" ~ 4,
    pan_exp_chg_pct == "41-50%" ~ 5,
    pan_exp_chg_pct == "51-60%" ~ 6,
    pan_exp_chg_pct == "61-70%" ~ 7,
    pan_exp_chg_pct == "71-80%" ~ 8,
    pan_exp_chg_pct == "81-90%" ~ 9,
    pan_exp_chg_pct == "91-100%" ~ 10,
    pan_exp_chg_pct == "No change" ~ 0,
    TRUE ~ as.numeric(pan_exp_chg_pct)  
  ))


reddit_finance_sub$pan_exp_chg_pct <- as.numeric(reddit_finance_sub$pan_exp_chg_pct)

# 6. pan_fi_chg
# As a result of the pandemic, did your FI (financially independent) number...

reddit_finance_sub$pan_fi_chg <- factor(reddit_finance_sub$pan_fi_chg, levels = c( "Decrease","No change", "Increase"))

# 7.pan_ret_date_chg
# As a result of the pandemic, did your planned RE (retirement) date...

reddit_finance_sub$pan_ret_date_chg <- factor(reddit_finance_sub$pan_ret_date_chg, levels = c( "Become later", "No change", "Become sooner"))

# 8. pan_financial_impact
#Overall, how would you characterize the pandemic's impact on your finances?

reddit_finance_sub$pan_financial_impact <- factor(reddit_finance_sub$pan_financial_impact, levels = c( "Negative", "Neutral", "Positive"))

# 9. political
# With which political party do you most closely identify? You do not need to be registered with a party to select it, answer based on your personal views.


reddit_finance_sub <- reddit_finance_sub %>%
  mutate(political_grouped = case_when(
    political %in% c("Democrat", "Libertarian Party", "Republican") ~ as.character(political),
    TRUE ~ "Other"
  ))

reddit_finance_sub$political_grouped <- as.factor(reddit_finance_sub$political_grouped)

# 10. race_eth
# What is your race/ethnicity? Select all that apply.

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(race_eth_grouped = case_when(
    race_eth %in% c("N/A", "NA", "Decline to state") ~ NA_character_,
    TRUE ~ as.character(race_eth)
  ))

reddit_finance_sub$race_eth_grouped <- as.factor(reddit_finance_sub$race_eth_grouped)

# 11. gender
# What is your gender?


reddit_finance_sub <- reddit_finance_sub %>%
  mutate(gender_grouped = case_when(
    tolower(gender) %in% c("decline to state", "n/a", "na") ~ NA_character_,
    tolower(gender) %in% c("non-binary") ~ "Non-Binary",
    TRUE ~ as.character(gender)
  ))

reddit_finance_sub$gender_grouped <- as.factor(reddit_finance_sub$gender_grouped)


# 12 age
# What is your age?
# For this variable, we apply the decision rule of replacing the intervals with their order number. For example, the interval "<18" will be replaced by 1, "18-23" by 2, and so on.
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(age = case_when(
    age %in% c("N/A", "NA") ~ NA_real_,
    age == "<18"   ~ 1,
    age == "18-23" ~ 2,
    age == "24-28" ~ 3,
    age == "29-33" ~ 4,
    age == "34-38" ~ 5,
    age == "39-43" ~ 6,
    age == "44-48" ~ 7,
    age == "49-53" ~ 8,
    age == "54-58" ~ 9,
    age == "59-63" ~ 10,
    TRUE ~ as.numeric(age)
  ))


reddit_finance_sub$age <- as.numeric(reddit_finance_sub$age)

# 13. edu
# What is the highest level of education you have completed?
#"For the case of the education variable, we grouped its values into broader categories."


reddit_finance_sub <- reddit_finance_sub %>%
  mutate(edu_grouped = case_when(
    edu %in% c("High School diploma / GED", "Less than high school", "Some high school") ~ "High School or Less",
    edu %in% c("Some college, no degree", "Trade School Degree") ~ "Some College or Trade School",
    edu %in% c("Associate's Degree", "Bachelor's Degree") ~ "Bachelor's or Associate's Degree",
    edu %in% c("Doctorate / Post Graduate", "Doctorate / Post-Graduate", "Graduate degree") ~ "Doctorate or Master's Degree",
    edu %in% c("N/A", "NA") ~ NA_character_,
    TRUE ~ as.character(edu)
  ))


edu_levels <- c("High School or Less", "Some College or Trade School", "Bachelor's or Associate's Degree",   "Doctorate or Master's Degree")
reddit_finance_sub$edu_grouped <- factor(reddit_finance_sub$edu_grouped, levels = edu_levels)


# 14. rel_status
# What is your relationship status?
 
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(rel_status_grouped = case_when(
    rel_status %in% c("Divorced", "Widowed", "Single, never married") ~ "Singles",
    rel_status %in% c("Divorced - Remarried", "In a relationship, but not married", "Married", "Widowed - Remarried") ~ "In a Relationship",
    rel_status %in% c("N/A", "NA") ~ NA_character_,
    TRUE ~ as.character(rel_status)
  )) %>%
  mutate(rel_status_grouped = as.factor(rel_status_grouped))

# 15. children
# Do you have children?


reddit_finance_sub <- reddit_finance_sub %>%
  mutate(children_grouped = case_when(
    children %in% c("Do not have children, and do not intend to", "Do not have children, but intend to") ~ "No Children",
    children %in% c("Have children") ~ "Have Children",
    children %in% c("N/A", "NA") ~ NA_character_,
    TRUE ~ as.character(children)
  )) %>%
  mutate(children_grouped = as.factor(children_grouped))

# 16. country
# What country are you in?

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(country_grouped = case_when(
    country %in% c("United States") ~ "United States",
    TRUE ~ "Other"
  )) %>%
  mutate(country_grouped = as.factor(country_grouped))

# 17. fin_indy
# Are you financially independent? Meaning you do not need to work for money, regardless of whether you work for money.


reddit_finance_sub$fin_indy <- ifelse(reddit_finance_sub$fin_indy == "Yes", 1, 0)

# 18. fin_indy_num
# At what amount invested will you consider yourself Financially Independent? (What is your FI number?)
reddit_finance_sub$fin_indy_num <- as.numeric(reddit_finance_sub$fin_indy_num)

# 19. fin_indy_pct
# What percent FI are you? (What percent of your FI number do you currently have?)
# While exploring this dataset, there are certain outlier values that appear to be errors, such as individuals who have 95000, 2.00E+05, 260000, 4800000 percent of their FI number. Therefore, we will replace these values with NA to avoid distorting our distribution of this variable

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(
    fin_indy_pct = as.numeric(fin_indy_pct),  
    fin_indy_pct = ifelse(fin_indy_pct > 1000, NA_real_, fin_indy_pct)  
  )


# 25 fin_indy_lvl
# Which of the following would you have considered yourself at the time you reached Financial Independence:

reddit_finance_sub <- reddit_finance_sub %>%
  mutate(
    fin_indy_lvl = case_when(
      fin_indy_lvl == "Don't understand" ~ NA_character_,
      TRUE ~ fin_indy_lvl
    ),
    fin_indy_lvl = as.factor(fin_indy_lvl)
  )

# 26. retire_age
# At what age do you intend to retire?
# For this variable, we apply the decision rule of replacing the intervals with their order number. For example, the interval "24-28" will be replaced by 1, "29-33" by 2, and so on.
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(retire_age = case_when(
    retire_age %in% c("Undecided") ~ NA_real_,
    retire_age == "24-28" ~ 1,
    retire_age == "29-33" ~ 2,
    retire_age == "34-38" ~ 3,
    retire_age == "39-43" ~ 4,
    retire_age == "44-48" ~ 5,
    retire_age == "49-53" ~ 6,
    retire_age == "54-58" ~ 7,
    retire_age == "59-63" ~ 8,
    retire_age == "64-68" ~ 9,
    retire_age == "69+"   ~ 10,
    TRUE ~ as.numeric(retire_age)
  ))


# 27. stp_whn_fin_indy
# Do you intend to stop working for money when you reach financial independence?

reddit_finance_sub$stp_whn_fin_indy <- as.factor(reddit_finance_sub$stp_whn_fin_indy)


# 28. industry
# Which of the following best describes the industry in which you currently or most recently work(ed)?
# After analyzing the distribution of industries in our dataset, we will group industries with lower frequency into 'Other Services'

reddit_finance_sub$industry_grouped <- as.factor(ifelse(
  reddit_finance_sub$industry %in% c(
    "Other Services", "Military", "Transportation",
    "Natural Resources", "Arts, Entertainment, Recreation",
    "Utilities", "Trade, transportation, utilities",
    "Agriculture, Forestry, Fishing", "Media", "Real Estate",
    "Leisure & Hospitality", "NA"
  ),
  "Other Services",
  reddit_finance_sub$industry
))


# 29. employer
# Which of the following best describes your current or most recent employer?
reddit_finance_sub$employer <- as.factor(reddit_finance_sub$employer)

# 30 role
# Which of the following best describes your current or most recent job role?
reddit_finance_sub$role <- as.factor(reddit_finance_sub$role)

# 31 ft_status
# What is your current employment status? - Full Time

reddit_finance_sub$ft_status <- ifelse(
  reddit_finance_sub$ft_status %in% c("N/A", "NA", "For an Organization (Employed), For Myself (Self-Employed)"),
  NA_character_,
  reddit_finance_sub$ft_status
)

reddit_finance_sub$ft_status <- as.factor(reddit_finance_sub$ft_status)


# 32. pt_status
# What is your current employment status? - Part Time, Regular

reddit_finance_sub$pt_status <- ifelse(
  reddit_finance_sub$pt_status %in% c("N/A", "NA", "For an Organization (Employed), For Myself (Self-Employed)"),
  NA_character_,
  reddit_finance_sub$pt_status
)

reddit_finance_sub$pt_status <- as.factor(reddit_finance_sub$pt_status)

# 33 gig_status
# What is your current employment status? -Side Gig, Intermittent

reddit_finance_sub$gig_status <- ifelse(
  reddit_finance_sub$gig_status %in% c("N/A", "NA", "For an Organization (Employed), For Myself (Self-Employed)"),
  NA_character_,
  ifelse(reddit_finance_sub$gig_status %in% c("For Myself (Self-Employed), N/A"), "For Myself (Self-Employed)", reddit_finance_sub$gig_status)
)

reddit_finance_sub$gig_status <- as.factor(reddit_finance_sub$gig_status)

# 34. ne_status
# What is your current employment status? -Not Employed

#It is decided to eliminate the variable 'ne_status' as it does not seem to make much sense. The question initially refers to people without employment, but later, in the responses, the type of employer is mentioned. Given this situation, it would be difficult to interpret.

# 35 edu_status
# What is your current educational status?

reddit_finance_sub$edu_status <- ifelse(reddit_finance_sub$edu_status %in% c("Full time student", "Part time student"), "Student", "Not a student")

reddit_finance_sub$edu_status <- as.factor(reddit_finance_sub$edu_status)


# 36 housing
# What is your current housing situation?


reddit_finance_sub <- reddit_finance_sub %>%
  mutate(housing_grouped = case_when(
    housing %in% c("Own",
                   "Own a house but currently living in a RV",
                   "Own a house that I rent out but I rent an apartment in another state",
                   "Rent and own",
                   "Rent my apartment, but own a home I rent out",
                   "Rent temporarily while building new home") ~ "Own",
    housing %in% c("Government housing",
                   "Homeless",
                   "House owned by parents",
                   "Houseless/nomadic - living in a car voluntarily",
                   "Live with family or friends",
                   "Live with family while building a house",
                   "Live with partner and split costs of owning",
                   "Live with partner in their house",
                   "Liveaboard boat",
                   "Living with parents",
                   "NA",
                   "Own and live in a van",
                   "Provided by employer or school",
                   "Rent") ~ "Not Own",
    TRUE ~ "Other"
  ))

reddit_finance_sub$housing_grouped <- as.factor(reddit_finance_sub$housing_grouped)

# 24, 37 - 65 
# For the income and expense variables, we will assume that missing values (NAs) indicate when a person has no expenses in that category. In such cases, we will consider these missing values to have a value of 0.

variables_to_replace <- c(
  "whn_fin_indy_num",
  "home_value",
  "brokerage_accts_tax",
  "retirement_accts_tax",
  "cash",
  "invst_accts",
  "spec_crypto",
  "invst_prop_bus_own",
  "other_val",
  "student_loans",
  "mortgage",
  "auto_loan",
  "credit_personal_loan",
  "medical_debt",
  "invst_prop_bus_own_debt",
  "other_debt",
  "2020_gross_inc",
  "2020_housing_exp",
  "2020_utilities_exp",
  "2020_transp_exp",
  "2020_necessities_exp",
  "2020_lux_exp",
  "2020_child_exp",
  "2020_debt_repay",
  "2020_invst_save",
  "2020_charity",
  "2020_healthcare_exp",
  "2020_taxes",
  "2020_edu_exp",
  "2020_other_exp"
)

# Replace NAs with 0 in the specified variables
reddit_finance_sub <- reddit_finance_sub %>%
  dplyr::mutate_at(vars(variables_to_replace), list(~replace_na(., 0)))


# Deleting outliers
#reddit_finance_sub <- reddit_finance_sub %>%
 # filter(abs(cash - mean(cash, na.rm = TRUE)) <= 3 * sd(cash, na.rm = TRUE))

#"Since there are many variables related to income or assets, debts, and expenses, we will create 3 variables that sum these individual items: total_debts, total_assets, and total_expenses

#total_assets
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(total_assets = rowSums(select(., 
    c("home_value","brokerage_accts_tax","retirement_accts_tax","cash","invst_accts","spec_crypto","invst_prop_bus_own","other_val"))))


#total_debts
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(total_debts = rowSums(select(., 
    c("student_loans","mortgage","auto_loan","credit_personal_loan","medical_debt","invst_prop_bus_own_debt","other_debt"))))

#total_expenses
reddit_finance_sub <- reddit_finance_sub %>%
  mutate(total_expenses = rowSums(select(., 
    c("2020_housing_exp", "2020_utilities_exp", "2020_transp_exp", "2020_necessities_exp", "2020_lux_exp", "2020_child_exp","2020_debt_repay", "2020_charity", "2020_healthcare_exp","2020_taxes", "2020_edu_exp", "2020_other_exp"))))

# We finally removed some previously transformed variables that we will not be using in their original form.
reddit_finance_sub <- reddit_finance_sub[, !(names(reddit_finance_sub) %in% c("political", "race_eth", "gender", "edu", "rel_status", "children","country", "industry","ne_status","housing","home_value","brokerage_accts_tax","retirement_accts_tax", "cash", "invst_accts","spec_crypto","invst_prop_bus_own","other_val","student_loans","mortgage","auto_loan","credit_personal_loan","medical_debt","invst_prop_bus_own_debt","other_debt","2020_housing_exp","2020_utilities_exp","2020_transp_exp","2020_necessities_exp","2020_lux_exp","2020_child_exp","2020_debt_repay","2020_charity","2020_healthcare_exp","2020_taxes","2020_edu_exp","2020_other_exp"))]
```

```{r}
#We are removing the NA records from the variables that we will use for our model.
columns <- c("age", "political_grouped", "edu_grouped", "housing_grouped", "total_debts", 
                      "total_assets", "total_expenses", "2020_gross_inc", "2020_invst_save")

reddit_finance_sub <- reddit_finance_sub[complete.cases(reddit_finance_sub[, columns]), ]

```

\vspace{0.5cm}

# **Overview**

The objective of this project is to address the research question:

**What factors contribute to an individual's perception of financial
independence?**

To accomplish this objective, we will perform logistic regression
analysis using a dataset that is a subset of the official results from
the 2020 Financial Independence Survey on Reddit. This subset includes
responses from individuals who represent themselves (excluding
contributions from other household members) and excludes responses from
retired individuals. The dataset encompasses **1,998 rows** and **65
variables**, covering information such as income contributors, the
financial impact of the pandemic, political affiliation, demographics,
details about financial independence, employment status, and various
financial aspects. The data has been sourced from Reddit. For additional
details regarding the data dictionary and source information, please
refer to the
[www.openintro.org](https://www.openintro.org/data/index.php?data=reddit_finance)
website.

In our original dataset, 147 individuals, which accounts for 8% of our
sample, reported considering themselves financially independent.
Therefore, we have unbalanced data, which we should take into
consideration in our future analyses.

\vspace{0.5cm}

# **Data cleaning**

In the data cleaning process, the following transformations were
implemented. For more details on the specific variables that experienced
these transformations, please refer to the appendix page.

-   The dependent variable **fin_indy** (Are you financially
    independent?) was transformed from a string variable to a binary
    numeric variable (0 and 1).

-   The variable **political** had 24 categories, which were reduced to
    4: 'Democrat,' 'Libertarian Party,' 'Republican,' and 'Other,'
    because the data had many categories with very few observations.
    Finally, it was transformed from string to factor.

-   The variable **age** was transformed from a string variable to a
    numeric variable to simplify analysis, with each category assigned a
    corresponding numeric value (1-8) to represent age brackets.

-   The variable **edu** had 11 categories, which were reduced to 4:
    "High School or Less", "Some College or Trade School", "Bachelor's
    or Associate's Degree", "Doctorate or Master's Degree", because the
    data had many categories with very few observations. Then, it was
    transformed from string to factor.

-   The variable **housing** (What is your current housing situation?)
    had 19 categories, which were reduced to 2: "Own", "Not Own",
    because the data had many categories with very few observations.
    Finally, it was transformed from string to factor.

-   Out of a total of 29 variables, which constituted numeric-type
    variables such as Children Expenses, Luxury Expenses, Transportation
    Expenses, Taxes, Medical Debt, etc., it was observed that they did
    not contain the number 0 but did have many NA values. The assumption
    was made that individuals who responded with NA in fields such as
    Children Expenses did so because they had no associated expenses in
    that category. For this reason, in these 29 variables, NA values
    were replaced with 0. The complete detail of the variables can be
    reviewed on the appendix page.

-   The variable **total_assets** was created, which is formed by the
    sum of 8 variables such as cash, investment accounts, crypto, etc.
    For more details, refer to the appendix.

-   The variable **total_debts** was created, which is formed by the sum
    of 7 variables such as student loans, mortgage, medical debt, etc.
    For more details, refer to the appendix.

-   The variable **total_expenses** was created, which is formed by the
    sum of 12 variables such as necessities expenses, children expenses,
    transportation expenses, etc.

-   Finally, observations with NA values for any of the above variables
    were removed, resulting in 11 observations out of the original total
    of 1,998, leaving us with 1,987 final observations.

# **Modeling**

\vspace{0.5cm}

## Logistic Regression Model

To address our research question, we will employ a Logistic Regression
Model. This model is suitable for this situation as it allows us to
examine how various predictor variables influence the probability of an
individual perceiving financial independence. Given that the variable of
interest is binary (yes/no), logistic regression will provide us with
estimations of log probabilities and coefficients that will aid in
understanding the direction and strength of the relationship between the
considered factors and the perception of financial independence.

\vspace{0.5cm}

## Variable Selection

After analyzing the variables provided in the dataset, those that could
be considered to influence the financial independence variable were
identified. The following variables were selected a priori as
predictors: **Age**, **Political** (With which political party do you
most closely identify? ), **Education** (What is the highest level of
education you have completed? ), **Housing** (What is your current
housing situation? **Rent**, Own...) , **Total Debts**, **Total
Assets**, **Total Expenses**, **2020 Gross Income**, **2020 Investments,
2020 savings**.

For the variable age, it is expected that as a person gets older, they
are more likely to have achieved greater economic stability and,
therefore, are more prone to being financially independent. Regarding
the political variable, it is anticipated that individuals belonging to
parties that oppose, for example, tax increases, see economic
independence as feasible through their own assets. It is expected that
with higher education, there is a greater likelihood of being
financially independent. Concerning the housing variable, it is expected
that owning a home will significantly impact the dependent variable. In
the case of the variables debt, assets, expenses, 2020 income, and 2020
investments, they are considered closely related to the dependent
variable, as the available wealth of an individual, i.e., income minus
expenses, assets minus liabilities, will determine whether they can
achieve economic independence or not.

\vspace{0.5cm}

# **Results**

```{r, results = FALSE}
logistic_mod <- glm(factor(fin_indy) ~ 
                     age+
                    political_grouped
                  + edu_grouped
                  + housing_grouped
                  + total_debts
                  + total_assets
                  + total_expenses
                  + `2020_gross_inc`
                  + `2020_invst_save`
                  ,data=reddit_finance_sub,
                  family="binomial")

summary(logistic_mod)
```

```{r}
rows_to_remove <-  c(168, 156, 167, 165)
reddit_finance_sub2 <- reddit_finance_sub[-rows_to_remove, ]

```

## **Model results**

```{r, results = TRUE}
logistic_mod2 <- glm(factor(fin_indy) ~ 
                    age
                  + political_grouped
                  + edu_grouped
                  + housing_grouped
                  + total_debts
                  + total_assets
                  + total_expenses
                  + `2020_gross_inc`
                  + `2020_invst_save`
                  ,data=reddit_finance_sub2,
                  family="binomial")

summary(logistic_mod2)
```

Age exhibits a positive association, indicating that as individuals age,
the likelihood of financial independence increases. Affiliation with the
Republican Party is associated with higher odds of financial
independence compared to the reference category. Education categories
show no statistically significant effects. Regarding housing,
individuals with "Other" housing arrangements are more likely to be
financially independent, while owning or renting does not show
significant effects. Notably, variables related to financial status,
such as total assets and total debts, demonstrate significant impacts,
with higher assets and lower debts associated with increased odds of
financial independence. The model's goodness of fit is evident from the
lower residual deviance compared to the null deviance. Overall, the
findings highlight the multifaceted nature of factors contributing to
financial independence.

A lower AIC value, specifically 820.88, implies a favorable balance
between the model's capacity to fit the data and its simplicity. This
suggests that the logistic regression model effectively captures the
data patterns considering its complexity.

## **ROC Curve**

```{r, fig.width=8, fig.height=6, results = TRUE}

roc1 <- roc(reddit_finance_sub$fin_indy, fitted(logistic_mod))
roc2 <- roc(reddit_finance_sub2$fin_indy, fitted(logistic_mod2))


# Create the ROC plot
plot(roc1, col = "blue", main = "ROC Curve Comparison")  # First ROC curve
lines(roc2, col = "red")  # Second ROC curve

# Add a legend
legend("bottomright", legend = c("Model 1", "Model 2"), col = c("blue", "red"), lwd = 2)

# Print the AUC values for both models
text(0.7, 0.3, paste("AUC Model 1 = ", round(roc1$auc, 4)), col = "blue")
text(0.7, 0.25, paste("AUC Model 2 = ", round(roc2$auc, 4)), col = "red")

# Print the best threshold for Model 2 with more decimals
best_threshold_model2 <- roc2$thresholds[which.max(roc2$sensitivities + roc2$specificities)]
text(0.7, 0.2, sprintf("Best Threshold Model 2 = %.4f", best_threshold_model2), col = "red")



```

Upon observing the ROC curve, it becomes evident that the model's
discriminative performance is robust, as indicated by the AUC value of
0.861. Notably, at the specific threshold of 0.067, a distinct point on
the curve is identified, emphasizing the tailored balance achieved
between the true positive and false positive rates. This nuanced insight
from the ROC analysis underscores the model's ability to effectively
classify instances with a refined understanding of the trade-off between
positive predictions and potential errors

Additionally, the confusion matrix was computed (see in the appendix).
The confusion matrix results reveal that, despite the low prevalence of
positive responses, the model achieves an accuracy of 79.82%.
Sensitivity and specificity metrics are crucial in this imbalanced
context, showcasing the model's ability to detect positive and negative
instances, respectively. The low precision (23.93%) suggests that
positive predictions should be approached with caution. Although the
Kappa coefficient implies some agreement beyond chance (0.2888), it is
vital to consider metrics adjusted for imbalance, such as sensitivity
and specificity, for a more comprehensive assessment of the model's
performance.

After generating our model, we can, for example, observe the behavior of
one of the variables that turned out to be most influential in
predicting our dependent variable, Total Assets. We can see how the
group of individuals who are financially independent indeed has
significantly higher assets.

```{r, fig.width=8, fig.height=5, results = TRUE}

reddit_finance_sub2$fin_indy <- as.factor(reddit_finance_sub2$fin_indy)

# Crear el boxplot sin outliers
boxplot(total_assets ~ fin_indy, data = reddit_finance_sub, 
        main = "Boxplot de Total Assets vs financial independence",
        xlab = "Fin Indy", ylab = "Total Assets",
        outline = FALSE)
```

# **Conclusion and Future work**

Given the previously results, the logistic regression model indicates a
positive association between age and financial independence, while
Republican Party affiliation, "Other" housing arrangements, and
favorable financial statuses significantly contribute. The model's
performance, reflected in the AUC of 0.861 and nuanced insights from the
ROC analysis, underscores its ability to discern positive instances.
However, the confusion matrix highlights the challenge of imbalanced
classes, urging caution regarding positive predictions. Despite an
accuracy of 79.82%, the evaluation of the model's practical utility
should prioritize precision and sensitivity.

The conducted analysis provides us with a model to identify factors
influencing people's perception of financial independence. While the
results make sense, it's crucial to consider study limitations. One
important aspect is the potential presence of confounding variables,
such as age, which could impact an individual's levels of assets,
liabilities, and income, factors theoretically influencing financial
independence. A future analysis could explore the results by removing
the confounding variable of age to more precisely assess the impact of
the specific factors considered in our model.

Furthermore, it would be relevant to conduct cross-validation analyses
to evaluate the model's generalization capacity to unseen data. This
helps verify if the model is robust and can be applied to new samples,
crucial for ensuring the reliability of conclusions and the practical
utility of the model in different contexts.

In conclusion, by addressing these considerations, we can enhance the
robustness and applicability of our model to understand the perception
of financial independence.

{{< pagebreak >}}

# Appendix

## Variable Transformation

+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| **Variable**   | **Transformation**                       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| political      | -   From 24 original categories, it was  |
|                |     grouped into 4:\                     |
|                |     "Democrat"\                          |
|                |     "Libertarian Party"\                 |
|                |     "Republican"\                        |
|                |     "Other" --\> ("American Solidarity   |
|                |     Party","Citizens Party of the United |
|                |     States","Communist Party             |
|                |     USA","Constitution Party","Freedom   |
|                |     Socialist Party","Green              |
|                |     Party","Humane Party","Independent   |
|                |     American Party","Legal Marijuana Now |
|                |     Party","N/A","NA","National          |
|                |     Socialist                            |
|                |     Movement","None","Objectivist        |
|                |     Party","Peace and Freedom            |
|                |     Party","Socialist Action","Socialist |
|                |     Alternative","Socialist Equality     |
|                |     Party","Socialist Party USA","United |
|                |     States Marijuana Party","United      |
|                |     States Pirate Party","Workers World  |
|                |     Party")                              |
|                |                                          |
|                | -   Transformed into a factor.           |
+----------------+------------------------------------------+
| age            | -   It was transformed from string to    |
|                |     numeric (1-10)                       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| edu            | -   From 11 original categories, it was  |
|                |     grouped into 4:\                     |
|                |     "High School or Less" --\> ("High    |
|                |     School diploma / GED", "Less than    |
|                |     high school", "Some high school")\   |
|                |     "Some College or Trade School" --\>  |
|                |     ("Some college, no degree", "Trade   |
|                |     School Degree")\                     |
|                |     "Bachelor's or Associate's Degree"   |
|                |     --\> ("Associate's Degree",          |
|                |     "Bachelor's Degree")\                |
|                |     "Doctorate or Master's Degree" --\>  |
|                |     ("Doctorate / Post Graduate",        |
|                |     "Doctorate / Post-Graduate",         |
|                |     "Graduate degree")                   |
|                | -   Transformed into a factor.           |
+----------------+------------------------------------------+
| fin_indy       | -   It was transformed from string to    |
|                |     binary                               |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| housing        | -   From the original 19 categories, it  |
|                |     was grouped into 2:\                 |
|                |     "Own" --\> (Own", "Own a house but   |
|                |     currently living in a RV", "Own a    |
|                |     house that I rent out but I rent an  |
|                |     apartment in another state", "Rent   |
|                |     and own", "Rent my apartment, but    |
|                |     own a home I rent out", "Rent        |
|                |     temporarily while building new       |
|                |     home")\                              |
|                |     "Not Own" --\> ("Government          |
|                |     housing", "Homeless", "House owned   |
|                |     by parents", "Houseless/nomadic -    |
|                |     living in a car voluntarily", "Live  |
|                |     with family or friends", "Live with  |
|                |     family while building a house",      |
|                |     "Live with partner and split costs   |
|                |     of owning", "Live with partner in    |
|                |     their house", "Liveaboard boat",     |
|                |     "Living with parents", "NA", "Own    |
|                |     and live in a van", "Provided by     |
|                |     employer or school", "Rent")         |
|                |                                          |
|                | -   Transformed into a factor.           |
+----------------+------------------------------------------+
| home_value     | -   NA replaced with 0                   |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| broke          | -   NA replaced with 0                   |
| rage_accts_tax |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| retire         | -   NA replaced with 0                   |
| ment_accts_tax |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| cash           | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| invst_accts    | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| spec_crypto    | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| invs           | -   NA replaced with 0                   |
| t_prop_bus_own |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| other_val      | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_assets variable         |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| student_loans  | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| mortgage       | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| auto_loan      | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| credi t\       | -   NA replaced with 0                   |
| \              |                                          |
| \              | -   sum in total_debts variable          |
| _personal_loan |                                          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| medical_debt   | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| invst_pr o     | -   NA replaced with 0                   |
| p_bus_own_debt |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| other_debt     | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_debts variable          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_gross_inc | -   NA replaced with 0                   |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2 0            | -   NA replaced with 0                   |
| 20_housing_exp |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 202 0\         | -   NA replaced with 0                   |
| \              |                                          |
| \              | -   sum in total_expenses variable       |
| _utilities_exp |                                          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2              | -   NA replaced with 0                   |
| 020_transp_exp |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_n         | -   NA replaced with 0                   |
| ecessities_exp |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_lux_exp   | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_child_exp | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2              | -   NA replaced with 0                   |
| 020_debt_repay |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2              | -   NA replaced with 0                   |
| 020_invst_save |                                          |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_charity   | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020\_         | -   NA replaced with 0                   |
| healthcare_exp |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_taxes     | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_edu_exp   | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+
| \vspace{0.3cm} | \vspace{0.3cm}                           |
|                |                                          |
| 2020_other_exp | -   NA replaced with 0                   |
|                |                                          |
|                | -   sum in total_expenses variable       |
+----------------+------------------------------------------+

{{< pagebreak >}}

## Cook's distance

### Model 1

```{r, results = TRUE}
plot(logistic_mod, which =4)
```

### Model 2

```{r, results = TRUE}
plot(logistic_mod2, which =4)
```

{{< pagebreak >}}

## Confusion Matrix

### Model 2

```{r, results = TRUE}
reddit_finance_sub2 <- reddit_finance_sub2 %>%
 mutate(fin_indy = ifelse(fin_indy == 1, "Yes", "No"))

confusionMatrix(factor(ifelse(fitted(logistic_mod2)>0.067,"Yes","No")),factor(reddit_finance_sub2$fin_indy),positive="Yes",mode="everything")

```

```{r}
print("hello")
```
